
This gist shows in two steps **how to tilt and stack maps using ggplot2** in order to create an image like this one:
[![enter image description here][1]][1]

Let's load the necessary libraries and data to use a reproducible example:
``` {r load libraries and data}
# load libraries
  library(rgeos)
  library(ggplot2)
  library(dplyr)
  library(RColorBrewer)
  library(UScensus2000tract)

# load data
  data("oregon.tract")

# plot Census Tract map
  plot(oregon.tract)
```

## 1st step: tilting maps
The 1st step is to tilt the map. This code comes from [Barry Rowlingson's](http://barry.rowlingson.com/) answer to a question on [gis.stackexchange](http://gis.stackexchange.com/questions/189490/plot-tilted-map-in-r). Thanks [Barry](https://twitter.com/geospacedman)! The idea here is to use a shear and a scale matrix to transform the coordinates of the spatial object once it has been fortified.

```{r tilt map}
# create a unique character ID value:
  oregon.tract$id=as.character(1:nrow(oregon.tract))

# Fortify on that ID and join attribute data:
  ofort <- fortify(oregon.tract,region="id")
  ofort <- left_join(ofort, oregon.tract@data, c("id"="id"))

# Shear/scale matrix [[2,1],[0,1]] obtained by some trial and error:
  sm <- matrix(c(2,1.2,0,1),2,2)

# Get transformed coordinates:
  xy <- as.matrix(ofort[,c("long","lat")]) %*% sm

# Add xy as extra columns in fortified data:
  ofort$x <- xy[,1]; ofort$y = xy[,2]

# Plot !
  ggplot(ofort, aes(x=x, y=y, group=id, fill=white)) + geom_polygon() + coord_fixed()
```
[![enter image description here][2]][2]


## 2nd step: stacking maps
The 2nd step is to stack different layers of the map. To do this, you just need to add another map layer and displace its `y` axis: e.g. `aes(x=x, y=y+5)` :

```{r Stack map}
    ggplot(data= ofort) + 
      geom_polygon( aes(x=x, y=y, group=id), fill= "white", color="gray30") + # layer 1
      geom_polygon( aes(x=x, y=y+5, group=id, fill=white)) +                  # layer 2
      geom_polygon( aes(x=x, y=y+10, group=id, fill=pop2000-white)) +         # layer 3
      theme(axis.text=element_blank(), axis.ticks=element_blank()) +
      scale_fill_distiller(palette = "RdBu", name = "Population") + # Use distiller variant with continous data
      annotate("text", x = -197, y = 45, size=5, color="gray35", label = "Census tracts") + # layer 1
      annotate("text", x = -197, y = 50, size=5, color="gray35", label = "White") +         # layer 2
      annotate("text", x = -197, y = 55, size=5, color="gray35", label = "Non-white") +     # layer 3
      coord_fixed()
```
[![enter image description here][1]][1]



  [1]: http://i.stack.imgur.com/3lbAT.png
  [2]: http://i.stack.imgur.com/kU1nZ.png
